Article Summaries

### 2025-01-28 articleSummaries.qmd
Updated article summaries

>>>>

[@asghar2024] The goal of the paper is to examine the survival and risk of death associated with Covid-19 in the hospitals of a Pakistani city. Specifically, the paper looks at asymptomatic vs symptomatic individuals and focuses on the effects of gender, age, and if the person was symptomatic to assist in determining an individual's survival if they had the virus. This paper and research is important as Covid-19 demonstrated that a virus can and will still spread massively throughout the world. Determining the chances of survival with the virus would have been beneficial in the height of the pandemic, but is also beneficial now as we are not going to stop living with the virus anytime soon. The data the paper uses is from the Directorate General Health Services in KP, Pakistan and contains the hospital attendance records from the first wave of the pandemic, specifically March 2020 to June 2020. Individuals age, gender, date of illness, dates of admission and discharge/deceased, and if the individual had symptoms or not when diagnosed with Covid-19. The study focuses on individuals who are younger than 100 and did not have missing values. They first examined the Kaplan-Meier curve for the survival probability from their data. This allowed them to determine that the conventional Cox proportional hazards model may not be best since patients would recover and continue to survive, but the Cox model assumes survival probability will eventually go to zero. As such, they also determined a second model using logisitic regression and the Weibull regression model. They used a mixture cure model and a non-mixture model for the Weibull regression model. Using the Cox proportional hazards model, with three covariates (age, gender, being asymptomatic), age and being asymptomatic were considered to be significantly associated with survival of Covid-19 patients. Patients who are older and were symptomatic had a higher risk of death. However, the model showed that gender was insignificant. For the Weibull mixture cure (which allows for a patients to be cured), older and symptomatic patients were again at an increased risk of death if the person was uncured. However, those patients did have a longer survival rate than the uncured asymptomatic patients. For the Weibull non-mixture cure model, age and being asymptomatic indicated an increased risk of death in the long-term survival, with the symptomatic indicating an increased risk in the short-term survival. Gender for both Weibull models was not considered significant.

[@jiang2024] The goal of the paper is to critique another paper on chest pain incidence and mortality in the US and provide the limits of the model used in that paper. Heart disease is the leading cause of death for the US according to the CDC. As such, research into the topic, especially attempting to predict the event of death should be thorough and well thought out. This paper address concerns for a research topic done. The study the paper focuses on found that there is a significant relationship between Triglyceride-Glucose and mortality. This was regardless of chest pain occurring within the patients. However, there were several limitations using the Cox proportional hazards model as the study did. The first limitation brought up is that the Cox model assumes constant hazard ratios over time, but that the effect of Triglyceride-Glucose could vary over time which would violate the assumption of proportional hazards ratios. That violation could lead to biased estimates and inaccurate conclusions. The Cox model also struggles with time-dependent variables. The study took into account body weight and blood glucose levels, both of which can (and will) change throughout time which could impact the true effect of Triglyceride-Glucose. The paper provides several other weaknesses of the original study. The paper ens by suggesting a better alternative approach would have been to use time-varying covariate models. This paper was helpful as it provided data driven examples of why Cox proportional hazards model may not be the best model. This will be helpful in guiding the research into datasets to model.
